---
output:
  rmarkdown::github_document
bibliography: "inst/references.bib"
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`JointXshift`

<!-- badges: start -->
[![R-CMD-check](https://github.com/blind-contours/JointXshift/workflows/R-CMD-check/badge.svg)](https://github.com/blind-contours/JointXshift/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/blind-contours/JointXshift/master.svg)](https://codecov.io/github/blind-contours/JointXshift?branch=master)
[![CRAN](https://www.r-pkg.org/badges/version/JointXshift)](https://www.r-pkg.org/pkg/JointXshift)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/JointXshift)](https://CRAN.R-project.org/package=JointXshift)
[![CRAN total downloads](http://cranlogs.r-pkg.org/badges/grand-total/JointXshift)](https://CRAN.R-project.org/package=JointXshift)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![MIT license](https://img.shields.io/badge/license-MIT-brightgreen.svg)](https://opensource.org/licenses/MIT)
<!-- [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4070042.svg)](https://doi.org/10.5281/zenodo.4070042) -->
<!-- [![DOI](https://joss.theoj.org/papers/10.21105/joss.02447/status.svg)](https://doi.org/10.21105/joss.02447) -->
<!-- badges: end -->

> Joint Shift Estimation using Stochastic Interventions
__Authors:__ [David McCoy](https://davidmccoy.org)

---

## What's `JointXshift`?


The `JointXshift` is a simple package that compares the expected outcome change under a joint stochastic shift intervention (the change in outcome compared to the observed outcome if all exposures were shifted by a fixed amount), to the additive expected outcome change, the additive impact of each exposure shifted individually. The idea is that if the joint shift is positively larger, this indicates synergistic relationships in the mixed exposure, if it is negative this indicates more antagonistic relationships exist. Of course, if there is no difference between the joint and additive this could indicate no interaction or interactions that are both synergistic and antagonistic and cancel each other out.

---

## Installation

*Note:* Because the `JointXshift` package (currently) depends on `sl3` that
allows ensemble machine learning to be used for nuisance parameter
estimation and `sl3` is not on CRAN the `JointXshift` package is not
available on CRAN and must be downloaded here.

There are many depedencies for `JointXshift` so it's easier to break up 
installation of the various packages to ensure proper installation. 

First install the basis estimators used in the data-adaptive
variable discovery of the exposure and covariate space: 

`JointXshift` uses the `sl3` package to build ensemble machine learners for each nuisance parameter. 

```{r sl3_devel,  eval = FALSE}
remotes::install_github("tlverse/sl3@devel")
```

Make sure `sl3` installs correctly then install `JointXshift`

```{r JointXshift_install,  eval = FALSE}
remotes::install_github("blind-contours/JointXshift@main")
```

---

## Example

To illustrate how `JointXshift` may be used to ascertain the effect of a mixed exposure, consider the following example:

```{r example, warning=FALSE}
library(JointXshift)
library(devtools)
library(kableExtra)
library(sl3)

seed <- 429153
set.seed(seed)
```

We will directly use synthetic data from the NIEHS used to test new mixture methods. This data has built in strong positive and negative marginal effects and certain interactions. Found here: https://github.com/niehs-prime/2015-NIEHS-MIxtures-Workshop

```{r NIEHS example}
data("NIEHS_data_1", package = "JointXshift")
```

```{r NIEH Nodes}
NIEHS_data_1$W <- rnorm(nrow(NIEHS_data_1), mean = 0, sd = 0.1)
w <- NIEHS_data_1[, c("W", "Z")]
a <- NIEHS_data_1[, c("X1", "X2", "X3", "X4", "X5", "X6", "X7")]
y <- NIEHS_data_1$Y

deltas <- list(
  "X1" = 1, "X2" = 1, "X3" = 1,
  "X4" = 1, "X5" = 1, "X6" = 1, "X7" = 1
)
head(NIEHS_data_1) %>%
  kbl(caption = "NIEHS Data") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


We will apply our method to see if there is largely more synergistic or antagonistic relationships in the data. 

```{r run JointXshift, eval = TRUE, message=FALSE, warning=FALSE}

ptm <- proc.time()
sim_results <- JointXshift(
  w = w,
  a = a,
  y = y,
  delta = deltas,
  n_folds = 3,
  num_cores = 6,
  outcome_type = "continuous",
  seed = seed)

proc.time() - ptm

## marginal effects
additive_results <- sim_results$`Additive Effects`
joint_results <- sim_results$`Joint Effects`
joint_vs_additive <- sim_results$`Joint vs Additive Effects`

```


The expected outcome change under additive assumptions is: 
```{r additive results}
additive_results
```


The expected outcome change under a joint shift is: 
```{r joint results}
joint_results
```

The difference is: 
```{r joint vs additive}
joint_vs_additive
```

This indicates that, compared to additive, the expected outcome change is -47.7, indicating possible more antagonistic relationships in the mixed exposure compared to synergistic. 

---

## Issues

If you encounter any bugs or have any specific feature requests, 
please [file an
issue](https://github.com/blind-contours/JointXshift/issues). Further details
on filing
issues are provided in our [contribution
guidelines](https://github.com/blind-contours/
JointXshift/main/contributing.md).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/blind-contours/JointXshift/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `JointXshift` R package, please cite the following:


---

## Related

* [R/`tmle3shift`](https://github.com/tlverse/tmle3shift) - An R package
  providing an independent implementation of the same core routines for the TML
  estimation procedure and statistical methodology as is made available here,
  through reliance on a unified interface for Targeted Learning provided by the
  [`tmle3`](https://github.com/tlverse/tmle3) engine of the [`tlverse`
  ecosystem](https://github.com/tlverse).

* [R/`medshift`](https://github.com/nhejazi/medshift) - An R package providing
  facilities to estimate the causal effect of stochastic treatment regimes in
  the mediation setting, including classical (IPW) and augmented double robust
  (one-step) estimators. This is an implementation of the methodology explored
  by @diaz2020causal.

* [R/`haldensify`](https://github.com/nhejazi/haldensify) - A minimal package
  for estimating the conditional density treatment mechanism component of this
  parameter based on using the [highly adaptive
  lasso](https://github.com/tlverse/hal9001) [@coyle-hal9001-rpkg;
  @hejazi2020hal9001-joss] in combination with a pooled hazard regression. This
  package implements a variant of the approach advocated by @diaz2011super.

---

## Funding

The development of this software was supported in part through 
NIH grant P42ES004705 from NIEHS

---

## License

&copy; 2020-2022 [David B. McCoy](https://davidmccoy.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License
Copyright (c) 2020-2022 David B. McCoy
Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:
The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.
THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References
