---
output:
  rmarkdown::github_document
bibliography: "inst/references.bib"
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`JointXshift`

<!-- badges: start -->
[![R-CMD-check](https://github.com/blind-contours/JointXshift/workflows/R-CMD-check/badge.svg)](https://github.com/blind-contours/JointXshift/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/blind-contours/JointXshift/master.svg)](https://codecov.io/github/blind-contours/JointXshift?branch=master)
[![CRAN](https://www.r-pkg.org/badges/version/JointXshift)](https://www.r-pkg.org/pkg/JointXshift)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/JointXshift)](https://CRAN.R-project.org/package=JointXshift)
[![CRAN total downloads](http://cranlogs.r-pkg.org/badges/grand-total/JointXshift)](https://CRAN.R-project.org/package=JointXshift)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![MIT license](https://img.shields.io/badge/license-MIT-brightgreen.svg)](https://opensource.org/licenses/MIT)
<!-- [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4070042.svg)](https://doi.org/10.5281/zenodo.4070042) -->
<!-- [![DOI](https://joss.theoj.org/papers/10.21105/joss.02447/status.svg)](https://doi.org/10.21105/joss.02447) -->
<!-- badges: end -->

> Joint Shift Estimation using Stochastic Interventions
__Authors:__ [David McCoy](https://davidmccoy.org)

---

## What's `JointXshift`?


The `JointXshift` R package offers an approach which identifies and estimates the impact interactions in a mixed exposure on an outcome. We define interaction as the counterfactual mean of the outcome under stochastic interventions of two exposures compared to the additive counterfactual mean of the two expowures intervened on indepdentently. These interventions or exposure changes depend on naturally observed values, as described in past literature [@diaz2012population; @haneuse2013estimation].

`JointXshift` builds on work described in [@mccoy2023semiparametric]. However instead of identifying interactions through an semi-parametric definition of an F-statistics and then estimating our interaction target parameter using CV-TMLE pooled across exposure sets, we provide a more streamlined approach. In this package, we identify interactions through g-computation first - evaluating the expected outcome under joint shift compared to the sum of individual shifts using Super Learner. We then rank these estimates as the highest sets of synergy and antagonism. We then use CV-TMLE and pool within the ranks. 


The package ensures robustness by employing a k-fold cross-validation framework. This framework helps in estimating a data-adaptive parameter, which is the stochastic shift target parameters for the exposure sets identified as having synergy or antagonism. The process begins by partitioning the data into parameter-generating and estimation samples. In the parameter-generating sample, we identify our ranks of antagonistic and synergistic exposure sets through a machine learning g-computation framework. In the estimation sample we then estimate our interaction target parameter using the doubly robust estimator TMLE to ensure asymptotic efficiency which allows us to construct confidence intervals for our estimates (unlike the g-comp method). 

By using JointXshift, users get access to a tool that offers both k-fold specific and aggregated results for the top synergistic and antagonistic relationships, ensuring that researchers can glean the most information from their data. For a more in-depth exploration, there's an accompanying vignette.

To utilize the package, users need to provide vectors for exposures, covariates, and outcomes. They also specify the respective $\delta$ for each exposure (indicating the degree of shift) and if this delta should be adaptive in response to positivity violations. The `top_n` parameter defines the top number of synergistic, antagonistic, positive and negative ranked impacts to estiamte. A detailed guide is provided in the vignette. With these inputs, `JointXshift` processes the data and delivers tables showcasing fold-specific results and aggregated outcomes, allowing users to glean insights effectively.

`JointXshift` also incorporates features from the `sl3` package [@coyle-sl3-rpkg], facilitating ensemble machine learning in the estimation process. If the user does not specify any stack parameters, `JointXshift` will automatically create an ensemble of machine learning algorithms that strike a balance between flexibility and computational efficiency.


---

## Installation

*Note:* Because the `JointXshift` package (currently) depends on `sl3` that
allows ensemble machine learning to be used for nuisance parameter
estimation and `sl3` is not on CRAN the `JointXshift` package is not
available on CRAN and must be downloaded here.

There are many depedencies for `JointXshift` so it's easier to break up 
installation of the various packages to ensure proper installation. 

First install the basis estimators used in the data-adaptive
variable discovery of the exposure and covariate space: 

`JointXshift` uses the `sl3` package to build ensemble machine learners for each nuisance parameter. 

```{r sl3_devel,  eval = FALSE}
remotes::install_github("tlverse/sl3@devel")
```

Make sure `sl3` installs correctly then install `JointXshift`

```{r JointXshift_install,  eval = FALSE}
remotes::install_github("blind-contours/JointXshift@main")
```

---

## Example

To illustrate how `JointXshift` may be used to ascertain the effect of a mixed exposure, consider the following example:

```{r example, warning=FALSE}
library(JointXshift)
library(devtools)
library(kableExtra)
library(sl3)

seed <- 429153
set.seed(seed)
```

We will directly use synthetic data from the NIEHS used to test new mixture methods. This data has built in strong positive and negative marginal effects and certain interactions. Found here: https://github.com/niehs-prime/2015-NIEHS-MIxtures-Workshop

```{r NIEHS example}
data("NIEHS_data_1", package = "JointXshift")
```

```{r NIEH Nodes}
NIEHS_data_1$W <- rnorm(nrow(NIEHS_data_1), mean = 0, sd = 0.1)
w <- NIEHS_data_1[, c("W", "Z")]
a <- NIEHS_data_1[, c("X1", "X2", "X3", "X4", "X5", "X6", "X7")]
y <- NIEHS_data_1$Y

deltas <- list(
  "X1" = 1, "X2" = 1, "X3" = 1,
  "X4" = 1, "X5" = 1, "X6" = 1, "X7" = 1
)
head(NIEHS_data_1) %>%
  kbl(caption = "NIEHS Data") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


Based on the data key, we expect X1 to have the strongest positive effect, X5 the strongest negative. So we would expect these to take the top ranks for these marginal associations. For interactions we expect to find interactions built into the data that are antagonistic and synergistic. The github page gives details on the types of interactions built into this synthetic data. The above image shows the interaction types that we might expect to find.


```{r run JointXshift, eval = TRUE, message=FALSE, warning=FALSE}

ptm <- proc.time()
sim_results <- JointXshift(
  w = w,
  a = a,
  y = y,
  delta = deltas,
  n_folds = 3,
  num_cores = 6,
  outcome_type = "continuous",
  seed = seed)

proc.time() - ptm

## marginal effects
additive_results <- sim_results$`Additive Effects`
joint_results <- sim_results$`Joint Effects`
joint_vs_additive <- sim_results$`Joint vs Additive Effects`

```


```{r additive results}
additive_results
```



```{r joint results}
joint_results
```


```{r joint vs additive}
joint_vs_additive
```


---

## Issues

If you encounter any bugs or have any specific feature requests, 
please [file an
issue](https://github.com/blind-contours/JointXshift/issues). Further details
on filing
issues are provided in our [contribution
guidelines](https://github.com/blind-contours/
JointXshift/main/contributing.md).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/blind-contours/JointXshift/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `JointXshift` R package, please cite the following:


---

## Related

* [R/`tmle3shift`](https://github.com/tlverse/tmle3shift) - An R package
  providing an independent implementation of the same core routines for the TML
  estimation procedure and statistical methodology as is made available here,
  through reliance on a unified interface for Targeted Learning provided by the
  [`tmle3`](https://github.com/tlverse/tmle3) engine of the [`tlverse`
  ecosystem](https://github.com/tlverse).

* [R/`medshift`](https://github.com/nhejazi/medshift) - An R package providing
  facilities to estimate the causal effect of stochastic treatment regimes in
  the mediation setting, including classical (IPW) and augmented double robust
  (one-step) estimators. This is an implementation of the methodology explored
  by @diaz2020causal.

* [R/`haldensify`](https://github.com/nhejazi/haldensify) - A minimal package
  for estimating the conditional density treatment mechanism component of this
  parameter based on using the [highly adaptive
  lasso](https://github.com/tlverse/hal9001) [@coyle-hal9001-rpkg;
  @hejazi2020hal9001-joss] in combination with a pooled hazard regression. This
  package implements a variant of the approach advocated by @diaz2011super.

---

## Funding

The development of this software was supported in part through 
NIH grant P42ES004705 from NIEHS

---

## License

&copy; 2020-2022 [David B. McCoy](https://davidmccoy.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License
Copyright (c) 2020-2022 David B. McCoy
Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:
The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.
THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References
